Measurement of ($\pi^-$-Ar) and ($K^+$-Ar) total hadronic cross sections in the LArIAT experiment

### 2018-05-28 AppendixB-TrackingStudies/B-Tracking.tex
s

\chapter{Tracking Studies}\label{ch:AppendixTrack}
In this section, we describe three studies. The first is a justification of the selection criteria for the beamline handshake with the TPC information.  We perform this study to boost  the correct identification of the particles in the TPC associated with the beamline information, while maintaining sufficient statistics for the cross section measurement.  The second study is an optimization of the tracking algorithm, with the scope of maximizing the identification of the hadronic interaction point inside the TPC.  These two studies are related, since the optimization of the tracking is performed on TPC tracks which have been matched to the wire chamber track; in turn, the tracking algorithm for TPC tracks determines the number of reconstructed tracks in each event used to try the matching with the wire chamber track. Starting with a sensible tracking reconstruction, we perform the WC2TPC matching optimization first, then the tracking optimization. The WC2TPC match purity and efficiency  are then calculated again with the optimized tracking.

\subsection{Study of WC to TPC Match}\label{ch:WC2TPCMatchOptimization}

Plots I want in this section:
\begin{enumerate}
\item WC2TPC MC DeltaX, DeltaY and $\alpha$
\end{enumerate}

Scope of this study is assessing the goodness of the wire chamber to TPC match on Monte Carlo and decide the selection values we will use on data. A word of caution is necessary here. With this study, we want to minimize pathologies associated with the presence of the primary hadron itself, e.g. the incorrect association between the beamline hadron and its decay products inside the TPC.  Assessing the contamination from pile-up\footnote{We remind the reader that the DDMC is a single particle Monte Carlo, where the beam pile up is not simulated.}, albeit related, is beyond the scope of this study.

@@

\subsection{Tracking Optimization}\label{ch:TrackOptimization}

>>>>

\chapter{Additional Tracking Studies for LArIAT Cross Section Analyses}\label{ch:AppendixTrack}
In this section, we describe two studies. The first is a justification of the selection criteria for the beamline handshake with the TPC information.  We perform this study to boost  the correct identification of the particles in the TPC associated with the beamline information, while maintaining sufficient statistics for the cross section measurement.  The second study is an optimization of the tracking algorithm, with the scope of maximizing the identification of the hadronic interaction point inside the TPC.  These two studies are related, since the optimization of the tracking is performed on TPC tracks which have been matched to the wire chamber track; in turn, the tracking algorithm for TPC tracks determines the number of reconstructed tracks in each event used to try the matching with the wire chamber track. Starting with a sensible tracking reconstruction, we perform the WC2TPC matching optimization first, then the tracking optimization. The WC2TPC match purity and efficiency  are then calculated again with the optimized tracking.

\subsection{Study of WC to TPC Match}

%Plots I want in this section:
%\begin{enumerate}
%\item WC2TPC MC DeltaX, DeltaY and $\alpha$
%\end{enumerate}

Scope of this study is assessing the performances of the WC2TPC match on Monte Carlo (see Section \ref{ch:WC2TPCMatchMethod} ) and decide the selection values we will use on data. A word of caution is necessary here. With this study, we want to minimize pathologies associated with the presence of the primary hadron itself, e.g. the incorrect association between the beamline hadron and its decay products inside the TPC.  Assessing the contamination from pile-up\footnote{We remind the reader that the DDMC is a single particle Monte Carlo, where the beam pile up is not simulated.}, albeit related, is beyond the scope of this study.

@@

\subsection{Tracking Optimization}\label{ch:TrackOptimization}
We perform an optimization of the clustering algorithm (see Section \ref{sec:SignalProc}) with the scope of maximizing the efficiency of finding the interaction point for the total hadronic cross section measurements. We define as the interaction point the most downstream point of a WC2TPC matched TPC tracks within the TPC fiducial volume. Since all the WC2TPC tracks are by definition beam particles, tracks travel from upstream to downstream in the TPC; thus, identifying the interaction point means to stop the tracking correctly.

TrajCluster is the package used to cluster hits in LArIAT; this package counts more than 20 tunable parameters. A standard method to develop clustering algorithms and checking their performances is to ``hand scan", which means recognizing the effect of parameters tuning by looking at a series of data event displays. Albeit we recognize the importance of hand scanning as a great diagnosis tool, we developed a fully automated optimization package which compares MC reconstructed information to MC truth.

We start by defining a figure of merit in order to discern what makes a parameter configuration better than an other. We chose the percentage of events whose reconstructed and true length differ less than 2 cm. We then identify the parameters in TrajCluster that are most important to  correctly stop  the tracking and an appropriate range of values for each of them. We chose to optimize the parameters that leverage on the angle between consecutive groups of hits, the number of hits use in the cluster fit and the average hit charge to stop the tracking. We define a configuration space with all  possible combination of values for the chosen parameters and we perform reconstruction one combination at a time: the combination with the highest figure of merit determines the optimized tracking reconstruction.

We chose construct the combination space using a total of 5 parameters, 3 values each and two iterations of the method (for a total of 486 combinations). We run the combinations on a sample of 100000 pion events. 
After the optimization, the most upstream point of the tracking is correctly identified 99.5\% of the times, the most downstream point is correctly identified 62.5\% of the times, the tracking stops short about 15\% of the times and misses the interaction point 22.5\% of the times. Hand scanning confirmed that the missed interaction points happen in the vast majority of cases for very shallow angles, as shown in the event display in Figure, or in the case of angles visible only in one projection plane. We also noticed that the premature stopping of the tracks is often related to the presence of delta rays parallel to the track. We see room of improvement, such as the delta ray removal and a forced track breaking in case of a kink present in a single plane, for a future analysis. \textcolor{red}{ADD evd}

The procedure behind this optimization package is virtually applicable to any LArSoft module where it is possible to define figure of merit.

### 2018-05-28 Chapter-8/8-Conclusions.tex
s

\chapter{Conclusions}\label{ch:Conclusions}
In the era of neutrino precision measurements, of huge liquid argon detectors and of massive amount of information from LArTPCs, a renew interest for an ancient measurement arises: the measurement of hadronic interactions with matter. With this work, we presented the first ever ($\pi^-$-Ar) and ($K^+$-Ar) total hadronic cross section measurements as a function of the hadron kinetic energy. These analyses are the first physics analyses developed by the LArIAT experiment.
Both the analysis follow a similar workflow and  they rely on beam line detector information as well as both calorimetry and tracking in the TPC.

In order to measure ($\pi^-$-Ar) total hadronic  argon cross sections, we start by selecting pion beamline candidates through a series of selections on the beamline and TPC information apt to maximize the number of pions in the selection over the number of muons and electrons. We use the LArIAT beamline MC to estimate the beam composition of the selected beamline candidare and we propagate the particle species to the LArAIT TPC constructing a properly weighted sample with the DDMC.

 The analyses start by identifying a sample of the hadron of interest in the beam line and assessing the beam line contaminations. It proceeds with tracking the hadron candidates in the TPC and measuring their kinetic energy at each point in the tracking: the fine sampling of an hadron in the TPC forms the set of ``incident" hadrons.  Then, the hadronic interaction point is identified and the raw cross section is calculated. Two corrections are then applied to the raw cross section -- a background subtractions and a correction for detector effects -- to obtain the true cross section measurement.\\

These analyses' work flow will serve as a basis for the future cross section measurements of pions and kaons in the exclusive channels.

>>>>

\chapter{Conclusions}\label{ch:Conclusions}
In the era of neutrino precision measurements, of huge liquid argon detectors and of massive amount of information from LArTPCs, a renewed interest for an ancient measurement arises: the measurement of hadronic interactions with matter. With this work, we presented the first ever ($\pi^-$-Ar) and ($K^+$-Ar) total hadronic cross section measurements as a function of the hadron kinetic energy. These analyses are the first physics analyses developed by the LArIAT experiment.
Both the analysis follow a similar workflow and  they rely on beam line detector information as well as both calorimetry and tracking in the TPC.

In order to measure ($\pi^-$-Ar) total hadronic  argon cross sections, we start by selecting pion beamline candidates through a series of selections on the beamline and TPC information apt to maximize the number of pions in the selection over the number of muons and electrons. We use the LArIAT beamline MC to estimate the beam composition of the selected beamline candidates and we propagate them to the LArAIT TPC constructing a properly weighted sample with the DDMC. We apply the thin slice method on the pion candidates and obtain the raw cross section measurement. From the simulated sample, we obtain two corrections accounting for the beamline background contamination and for detector effects. Finally, we apply the corrections to data and measure the true cross section.

In order to measure ($K^+$-Ar) total hadronic  argon cross sections, we follow a similar procedure, i.e. we apply the thin slice method on kaon candidates identified in the beamline to obtain the raw cross section. We apply a background correction and a correction for detector effects to the raw cross section.
The background correction accounts for the presence of secondary particles in both the interacting and incident histograms and for the presence of decay events in the interacting plot.

The final results for the ($\pi^-$-Ar) and ($K^+$-Ar) total hadronic cross section are shown side by side in figure \ref{fig:finalfinal}.

\begin{figure}[htb]
\centering
\includegraphics[width=0.48\textwidth]{Chapter-6/Images/TheMoneyPlot.pdf}
\includegraphics[width=0.48\textwidth]{Chapter-7/Images/TheMoneyPlotK.pdf}
\caption{\emph{Left:} ($\pi^-$-Ar) total hadronic cross section measurements in the 60A and 100A samples overlaid with the  Geant4 prediction (green). \emph{Right:} ($K^+$-Ar) total hadronic cross section for  scattering angles greater than 5$^\circ$ measured in the 60A sample, statistical uncertainty in black and systematic uncertainty in red. The Geant4 prediction for the total hadronic cross section for angle scattering greater than 5$^\circ$ is displayed in green. } 
\label{fig:finalfinal}
\end{figure}

These analyses' will serve as a basis for the future cross section measurements of pions and kaons for the exclusive channels in LArIAT.
